Survival Analysis

### 2014-04-03 presentation.Rnw
Finish models
they're really quite simple!
I call this done-done!

\begin{frame}{Models}
(Cox) Proportional Hazards Model
% semi-parametric
% hazards function given covariates (factors)
% H_0: all \Beta_i = 0

$$\lambda(t | \mathbf{x}) = \lambda_{0}(t) \exp \left( \mathbf{x' \beta} \right)$$

Accelerated Failure Time Model
% parametric, specifies log(T)
%   exponential
%   weibull
%   lognormal
%   log-log
%   gamma

$$\log T = \mathbf{x' \beta} + \mathbf{\epsilon}$$

Exponential/Weibull Model

$$\lambda(t, \mathbf{x}) = \lambda_{0} \exp \left( \mathbf{x' \beta} \right)$$
\end{frame}

%%% http://www2.sas.com/proceedings/sugi26/p244-26.pdf
%%% http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3227332/
%%% http://www.jehps.net/juin2009/Aalenetal.pdf
\end{document}

>>>>

\begin{frame}{Nelson-Aalen Estimator for $H(t)$}
\begin{eqnarray*}
                             \hat H(0) & = & 0 \\
                             \hat H(t) & = & \sum_{t_i = 0}^{t} \frac{d_i}{n_i} \\
                                       & = & \frac{d_i}{n_i} + H(t_{i - 1}) \\
  \mathrm{Var}\left[ \hat H(t) \right] & = &
  \sum_{t_i = 0}^{t} \frac{(n_i - d_i) d_i}{(n_i - 1) n_i^2}
\end{eqnarray*}
\end{frame}

\begin{frame}{Models}
(Cox) Proportional Hazards Model
% examines cumulative hazards
% vary lambda_0: different distributions
% eg. lambda_0 ~ constant => exponential
% eg. lambda_0 ~ weibull => weibull
$$\lambda(t | \mathbf{x}) = \lambda_{0}(t) \exp \left( \mathbf{x' \beta} \right)$$

Accelerated Failure Time Model
% examines survival time
% vary e: different distributions
% eg. e_i ~ normal => log-normal
$$\log T = \mathbf{x' \beta} + \mathbf{\epsilon}$$
\end{frame}

\begin{frame}[fragile]{Confirmatory Statistics}
Are all the hazard curves the same?

\begin{eqnarray*}
  H_0 & : & \text{all } H_i = H_j \\
  H_a & : & \text{some } H_i \neq H_j
\end{eqnarray*}

\scriptsize
<<bfeed-test-correspondence>>=
survdiff(Surv(duration, delta) ~ poverty, data = bfeed.short)
@
\end{frame}

\begin{frame}[fragile]{Confirmatory Statistics}
Are two survival curves the same?

\begin{eqnarray*}
  H_0 & : & H_{s \leq 12} = H_{s > 12} \\
  H_0 & : & H_{s \leq 12} \neq H_{s > 12}
\end{eqnarray*}

\tiny
<<bfeed-test-single>>=
summary(coxph(Surv(duration, delta) ~ yschool > 12, data = bfeed.short))
@
\end{frame}

%%% http://data.princeton.edu/wws509/notes/c7.pdf
%%% http://www2.sas.com/proceedings/sugi26/p244-26.pdf
%%% http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3227332/
%%% http://www.jehps.net/juin2009/Aalenetal.pdf
\end{document}
